---
title: ""
author: "CWJvanBlommestwin"
date: "27022025"
output: html_document: keep_md: true
---

# Introduction

This report analyzes activity monitoring data, including total steps per day, average steps per interval, and differences between weekday and weekend activity patterns.

```{r setup, include=FALSE}
# Load necessary libraries
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(gridExtra)
```

# Data Preparation

```{r load-data}
# Read data
data <- read.csv("activity.csv", stringsAsFactors = FALSE)

# Convert 'date' to Date format
data$date <- as_date(data$date)
```

## Summary of Missing Values

```{r missing-values}
# Count missing values
missing_values <- sapply(data, function(x) sum(is.na(x)))
missing_values
```

# Step Count Analysis

## Total Steps per Day

```{r total-steps-day}
# Aggregate total steps per day
agg_data <- data %>%
  group_by(date) %>%
  summarise(steps_per_day = sum(steps, na.rm = TRUE))

# Plot histogram of total steps per day
ggplot(agg_data) +
  geom_histogram(aes(steps_per_day), fill = "white", color = "blue", bins = 30) +
  labs(title = "Total Steps per Day", x = "Total Steps", y = "Number of Days")
```

## Mean and Median Steps per Day

```{r mean-median-steps}
mean_steps_day <- mean(agg_data$steps_per_day, na.rm = TRUE)
med_steps_day <- median(agg_data$steps_per_day, na.rm = TRUE)

# Print results
mean_steps_day
med_steps_day
```

# Activity Patterns Across Intervals

```{r avg-steps-interval}
# Aggregate average steps per interval
agg_interval <- data %>%
  group_by(interval) %>%
  summarise(avg_steps_per_interval = mean(steps, na.rm = TRUE))

# Plot average steps per interval
ggplot(agg_interval) +
  geom_line(aes(interval, avg_steps_per_interval), lwd = 1.1, color = "blue") +
  labs(title = "Average Steps per Interval", x = "Interval", y = "Average Steps")
```

## Interval with Maximum Average Steps

```{r max-interval}
max_interval <- filter(agg_interval, avg_steps_per_interval == max(avg_steps_per_interval))
max_interval
```

# Imputing Missing Values

```{r impute-missing-values}
# Fill missing values using mean steps per interval
data_filled <- data %>%
  left_join(agg_interval, by = "interval") %>%
  mutate(steps = ifelse(is.na(steps), avg_steps_per_interval, steps))
```

## Total Steps per Day (After Imputation)

```{r total-steps-imputed}
agg_filled <- data_filled %>%
  group_by(date) %>%
  summarise(steps_per_day = sum(steps))

# Plot histogram of total steps per day after imputation
ggplot(agg_filled) +
  geom_histogram(aes(steps_per_day), fill = "white", color = "blue", bins = 30) +
  labs(title = "Total Steps per Day (Imputed)", x = "Total Steps", y = "Number of Days")
```

# Weekday vs Weekend Activity

```{r weekday-weekend}
# Add weekday information
data_filled <- data_filled %>%
  mutate(week_day = wday(date, label = TRUE))

# Separate weekdays and weekends
weekdays_data <- filter(data_filled, week_day %in% c("Mon", "Tue", "Wed", "Thu", "Fri"))
weekends_data <- filter(data_filled, week_day %in% c("Sat", "Sun"))

# Aggregate average steps per interval for weekdays and weekends
agg_weekdays <- weekdays_data %>%
  group_by(interval) %>%
  summarise(avg_steps_per_interval = mean(steps, na.rm = TRUE))

agg_weekends <- weekends_data %>%
  group_by(interval) %>%
  summarise(avg_steps_per_interval = mean(steps, na.rm = TRUE))

# Plot weekday and weekend patterns
plot_weekdays <- ggplot(agg_weekdays) +
  geom_line(aes(interval, avg_steps_per_interval), lwd = 1.1, color = "blue") +
  labs(title = "Weekday Activity", x = "Interval", y = "Avg Steps") +
  scale_y_continuous(limits = c(0, 220))

plot_weekends <- ggplot(agg_weekends) +
  geom_line(aes(interval, avg_steps_per_interval), lwd = 1.1, color = "blue") +
  labs(title = "Weekend Activity", x = "Interval", y = "Avg Steps") +
  scale_y_continuous(limits = c(0, 220))

grid.arrange(plot_weekdays, plot_weekends, nrow = 2)
```

# Conclusion

-   The mean number of steps per day is `r mean_steps_day`, and the median is `r med_steps_day`.
-   The highest average steps occur at interval `r max_interval$interval`, with `r max_interval$avg_steps_per_interval` steps.
-   Imputing missing values changed the step distribution.
-   Step activity patterns differ between weekdays and weekends.

This report provides insight into daily step counts and interval-based movement patterns, helping to understand trends in physical activity.
